---
params:
  county: "Salt Lake"
  data_dir: "data/remotes"
format:
  typst:
    fontsize: 8pt
    keep-typ: false
---

```{=typst}
#set text(font: "Poppins", size: 8pt, number-type: "lining", number-width: "tabular")
#show table.cell: set block(breakable: false)
```

```{r}
#| label: setup
#| include: false
library(dplyr)
library(yaml)

settings <- yaml::read_yaml("config/settings.yml")
report_data <- readRDS(file.path(params$data_dir, "report_data.rds"))

county_data <- dplyr::filter(report_data, county == params$county) |>
  dplyr::arrange(GEOID)
```

```{r}
#| label: table
#| results: asis
#| echo: false
#| warning: false
# ── Palette ───────────────────────────────────────────────────────────────────
teal <- "#003B4F"
teal_mid <- "#0E5C72"
teal_soft <- "#3E7E91"
sky <- "#7EC8E3"
amber <- "#C77B30"
key_tint <- "#F1F7F8"
line_cool <- "#CADDE2"
hair <- "#EEF3F4"
ink <- "#16272D"
body_c <- "#3C4A4F"
muted_c <- "#7A848A"
track <- "#E4ECEE"

# ── Helpers ───────────────────────────────────────────────────────────────────
fmt_val <- function(x, digits = 1, suffix = "", big_mark = "") {
  if (length(x) == 0 || is.na(x)) {
    return("—")
  }
  if (is.numeric(x) && !is.finite(x)) {
    return("—")
  }
  rounded <- round(x, digits)
  if (digits == 0 && nchar(big_mark) > 0) {
    return(paste0(
      format(as.integer(rounded), big.mark = big_mark, scientific = FALSE),
      suffix
    ))
  }
  paste0(format(rounded, scientific = FALSE), suffix)
}

col_val <- function(data, col) {
  if (col %in% names(data)) data[[col]] else NA_real_
}

oz_color <- function(cls) {
  dplyr::case_when(
    grepl("Likely to attract", cls, ignore.case = TRUE) &
      !grepl("More|Less|Un", cls, ignore.case = TRUE) ~ "#15705A",
    grepl("More likely", cls, ignore.case = TRUE) ~ "#2F8FB0",
    grepl("Less likely", cls, ignore.case = TRUE) ~ "#D98A3D",
    grepl("Unlikely|not likely", cls, ignore.case = TRUE) ~ "#B4532E",
    TRUE ~ "#9AA3A8"
  )
}

oz_class_short <- function(cls) {
  dplyr::case_when(
    grepl("Likely to attract", cls, ignore.case = TRUE) &
      !grepl("More|Less|Un", cls, ignore.case = TRUE) ~ "Likely",
    grepl("More likely", cls, ignore.case = TRUE) ~ "More Likely",
    grepl("Less likely", cls, ignore.case = TRUE) ~ "Less Likely",
    grepl("Unlikely|not likely", cls, ignore.case = TRUE) ~ "Unlikely",
    TRUE ~ "Unknown"
  )
}

# In-cell bar: number + thin bar below; widths use format() not sprintf to avoid % issues
cell_bar <- function(num_str, bar_pct_val, bar_color) {
  track_bar <- paste0(
    '#block(width: 100%, height: 3pt, fill: rgb("',
    track,
    '"), radius: 1.5pt)'
  )
  if (
    is.null(bar_pct_val) ||
      is.na(bar_pct_val) ||
      !is.finite(as.numeric(bar_pct_val))
  ) {
    return(paste0(
      '#text(size: 9pt, weight: "semibold")[',
      num_str,
      ']#v(2pt)',
      track_bar,
      '[]'
    ))
  }
  fp <- min(max(as.numeric(bar_pct_val), 0), 100)
  fpp <- paste0(format(round(fp, 1), nsmall = 1), "%")
  paste0(
    '#text(size: 9pt, weight: "semibold")[',
    num_str,
    ']',
    '#v(2pt)',
    track_bar,
    '[#block(width: ',
    fpp,
    ', height: 3pt, fill: rgb("',
    bar_color,
    '"), radius: 1.5pt)[]]'
  )
}

# Short tract label from GEOID last 6 digits
tract_label <- function(geoid) {
  sfx <- substr(as.character(geoid), 6, 11)
  base <- as.integer(substr(sfx, 1, 4))
  dec <- substr(sfx, 5, 6)
  if (dec == "00") paste0("Tract ", base) else paste0("Tract ", base, ".", dec)
}

# ── Column maxima for relative bar scaling ────────────────────────────────────
safe_max <- function(x) {
  v <- suppressWarnings(max(x, na.rm = TRUE))
  if (!is.finite(v) || v == 0) NA_real_ else v
}
col_max <- list(
  oz1_pct = safe_max(county_data$oz1_pct),
  muc = safe_max(county_data$wc_muc_pct),
  ato = safe_max(county_data$ato_composite),
  sap = safe_max(county_data$sap_planned_hh_per_acre),
  eru = safe_max(county_data$eru_per_acre)
)
rel_pct <- function(val, mx) {
  if (is.na(val) || !is.finite(val) || is.na(mx)) {
    NA_real_
  } else {
    min(val / mx * 100, 100)
  }
}

n_tracts <- nrow(county_data)
county_name <- params$county

# ── OZ legend swatches (pure paste0, no %%) ──────────────────────────────────
legend_swatches <- paste0(
  '#box(fill: rgb("#2F8FB0"), radius: 2pt, inset: (x: 5pt, y: 2pt))',
  '[#text(fill: white, size: 7pt)[More Likely]]#h(4pt)',
  '#box(fill: rgb("#15705A"), radius: 2pt, inset: (x: 5pt, y: 2pt))',
  '[#text(fill: white, size: 7pt)[Likely]]#h(4pt)',
  '#box(fill: rgb("#D98A3D"), radius: 2pt, inset: (x: 5pt, y: 2pt))',
  '[#text(fill: white, size: 7pt)[Less Likely]]',
  '#h(1fr)',
  '#text(fill: rgb("',
  muted_c,
  '"), size: 7pt)',
  '[Bars show value relative to column maximum · per-tract maps on detail sheets.]'
)

# ── Page setup: running header + footer via #set page ────────────────────────
# NOTE: all % in paste0 are literal — no %% needed (that's only for sprintf)
hdr_block <- paste0(
  '#block(fill: rgb("',
  teal,
  '"), width: 100%, inset: (x: 0.15in, y: 0.14in), radius: 6pt)',
  '[',
  # Right panel: placed at right + horizon so it centers vertically within the block
  # (block height is set by the left content; #place uses that height for centering)
  '#place(right + horizon)',
  '[#grid(columns: (auto, auto, auto), gutter: 10pt, align: horizon,',
  # Sub-group 1: tract count (white) + ELIGIBLE TRACTS label (sky)
  '[#stack(dir: ttb, spacing: 5pt,',
  '[#text(fill: white, size: 20pt, weight: "bold")[',
  n_tracts,
  ']],',
  '[#text(fill: rgb("',
  sky,
  '"), size: 5pt, weight: "semibold", tracking: 0.5pt)[ELIGIBLE TRACTS]]',
  ')],',
  # Thin vertical rule separator
  '[#align(horizon)[#block(width: 0.5pt, height: 0.31in, fill: rgb("',
  sky,
  '").transparentize(35%))[]]]',
  ',',
  # Sub-group 2: current page (white large) + /total (sky smaller) on one line, PAGE label below
  '[#stack(dir: ttb, spacing: 5pt,',
  '[#context [#text(fill: white, weight: "bold", size: 12pt)[#counter(page).display()]',
  '#text(fill: rgb("',
  sky,
  '"), weight: "semibold", size: 12pt)[/#counter(page).final().first()]]],',
  '[#text(fill: rgb("',
  sky,
  '"), size: 5pt, weight: "semibold", tracking: 0.5pt)[PAGE]]',
  ')]',
  ')]', # closes inner right grid ) + #place content ]
  # Left content: badge + county name/subtitle — determines block height
  '#grid(columns: (auto, 1fr), gutter: 0.07in, align: left + horizon,',
  '[#rect(fill: none, stroke: (paint: rgb("',
  sky,
  '"), thickness: 0.5pt), inset: (x: 6pt, y: 5pt), radius: 2pt)',
  '[#align(center)[#text(fill: rgb("',
  sky,
  '"), weight: "bold", size: 6pt)[WFRC#linebreak()MAG]]]],',
  '[#stack(dir: ttb, spacing: 10pt,',
  '[#text(fill: white, weight: "bold", size: 16pt)[',
  county_name,
  ' County]],',
  '[#text(fill: rgb("',
  sky,
  '"), size: 8pt)[Opportunity Zone 2.0 — Eligible Census Tracts]]',
  ')]',
  ')]' # closes left grid ) + outer block ]
)

legend_block <- paste0(
  '#block(width: 100%, above: 10pt, inset: (x: 0.15in, y: 2pt))',
  '[#box(inset: (y: 2pt))[#text(fill: rgb("',
  amber,
  '"), weight: "bold", size: 6pt, tracking: 1pt)[INVESTMENT LIKELIHOOD]]',
  '#h(6pt)',
  legend_swatches,
  ']'
)

hdr_content <- paste0(
  '#block(width: 100%, height: 0.28in)[]\n',
  hdr_block,
  '\n',
  legend_block,
  '\n',
  '#block(width: 100%, height: 8pt, above: 0pt, below: 0pt)[]\n',
  '#line(length: 100%, stroke: (paint: rgb("',
  line_cool,
  '"), thickness: 0.5pt))\n'
)

footer_content <- paste0(
  '#line(length: 100%, stroke: (paint: rgb("',
  line_cool,
  '"), thickness: 0.5pt))',
  '#v(1pt)',
  '#grid(columns: (1fr, auto), gutter: 4pt,',
  '[#text(fill: rgb("',
  muted_c,
  '"), size: 6pt)',
  '[Sources: Federal OZ 2.0 eligibility, Urban Institute OZ Tool, WFRC/MAG regional data. Growth projections 2027-2037.]],',
  '[#text(fill: rgb("',
  muted_c,
  '"), size: 6pt)',
  '[ATO = Access to Opportunities · ERU = Equiv. Residential Unit · SAP = Station Area Plan · MUorC = Metro/Urban/City Center]])'
)

# Emit the #set page() block
cat("``` {=typst}\n")
cat(paste0(
  '#set page(\n',
  '  flipped: true,\n',
  '  margin: (x: 0.4in, top: 1.32in, bottom: 0.5in),\n',
  '  header-ascent: 0pt,\n',
  '  footer-descent: 0pt,\n',
  '  header: [',
  hdr_content,
  '],\n',
  '  footer: [',
  footer_content,
  '],\n',
  ')\n'
))
cat("```\n\n")

# ── Two-tier table ────────────────────────────────────────────────────────────
cat("``` {=typst}\n")

# Resolve a color token: named Typst colors pass through; hex strings get rgb()
typst_color <- function(c) {
  if (c %in% c("white", "black")) c else paste0('rgb("', c, '")')
}

# Super-header cell builder
grp_hdr <- function(label, colspan, bg, text_clr = "white") {
  paste0(
    'table.cell(colspan: ',
    colspan,
    ', fill: ',
    typst_color(bg),
    ', align: center + bottom)',
    '[#text(fill: ',
    typst_color(text_clr),
    ', weight: "bold", size: 6pt, tracking: 0.5pt)[',
    toupper(label),
    ']]'
  )
}

# Column header cell builder
col_hdr <- function(label, bg = teal, fg = "white") {
  paste0(
    'table.cell(fill: ',
    typst_color(bg),
    ', align: center + bottom)',
    '[#text(fill: ',
    typst_color(fg),
    ', weight: "bold", size: 7.5pt)[',
    label,
    ']]'
  )
}

# Table declaration: 10 columns
cat(paste0(
  '#table(\n',
  '  columns: (0.27in, 1.55in, 0.85in, 0.85in, 0.66in, 1fr, 1fr, 1fr, 1fr, 1.12in),\n',
  '  stroke: none,\n',
  '  inset: (x: 5pt, y: 5pt),\n',
  '  fill: (col, row) => {\n',
  '    if row > 1 {\n',
  '      if col >= 5 and col <= 8 {\n',
  '        if calc.odd(row) { rgb("',
  key_tint,
  '") } else { white }\n',
  '      } else {\n',
  '        if calc.odd(row) { rgb("',
  hair,
  '") } else { white }\n',
  '      }\n',
  '    }\n',
  '  },\n',
  '  align: (col, row) => {\n',
  '    if row <= 1 { center + bottom }\n',
  '    else if col == 0 { center + top }\n',
  '    else if col == 1 { left + top }\n',
  '    else if col == 9 { center + top }\n',
  '    else { right + top }\n',
  '  },\n'
))

# Two-tier table.header
cat('  table.header(\n')
# Row 0: super-headers
cat(paste0('    ', grp_hdr("", 1, teal), ',\n'))
cat(paste0('    ', grp_hdr("Identity", 1, teal), ',\n'))
cat(paste0('    ', grp_hdr("Profile", 3, teal), ',\n'))
# Four Key Measures super-header (amber text on key-tint bg)
cat(paste0(
  '    table.cell(colspan: 4, fill: rgb("',
  key_tint,
  '"), align: center + bottom)',
  '[#text(fill: rgb("',
  amber,
  '"), weight: "bold", size: 6pt, tracking: 0.5pt)',
  '[FOUR KEY MEASURES]],\n'
))
cat(paste0('    ', grp_hdr("", 1, teal), ',\n'))
# Row 1: column headers
cat(paste0('    ', col_hdr("\\#"), ',\n'))
cat(paste0('    ', col_hdr("Tract"), ',\n'))
cat(paste0('    ', col_hdr("Pop · HH"), ',\n'))
cat(paste0('    ', col_hdr("Acres · Dev"), ',\n'))
cat(paste0('    ', col_hdr("OZ 1.0 %"), ',\n'))
cat(paste0(
  '    ',
  col_hdr("Wasatch Choice#linebreak()Centers (% Area)", key_tint, ink),
  ',\n'
))
cat(paste0(
  '    ',
  col_hdr("Accessibility#linebreak()Score /100", key_tint, ink),
  ',\n'
))
cat(paste0(
  '    ',
  col_hdr("Station Area Plan#linebreak()HH/ac", key_tint, ink),
  ',\n'
))
cat(paste0(
  '    ',
  col_hdr("Projected 2027–37#linebreak()Growth ERU/ac", key_tint, ink),
  ',\n'
))
cat(paste0('    ', col_hdr("Investment#linebreak()Likelihood"), ',\n'))
cat('  ),\n')

# ── Data rows ─────────────────────────────────────────────────────────────────
for (i in seq_len(nrow(county_data))) {
  tract <- county_data[i, ]
  geoid <- as.character(tract$GEOID)
  g <- function(col) col_val(tract, col)
  fv <- function(col, d = 1, s = "", m = "") fmt_val(g(col), d, s, m)

  # Col 1: running index (zero-padded)
  cat(paste0(
    '  [#text(fill: rgb("',
    muted_c,
    '"), size: 7pt)[',
    sprintf("%02d", i),
    ']],\n'
  ))

  # Col 2: tract label + GEOID (mono muted)
  pfx <- substr(geoid, 1, 5)
  sfx <- substr(geoid, 6, 11)
  cat(paste0(
    '  [#text(weight: "semibold", size: 8pt)[',
    tract_label(geoid),
    ']',
    '#linebreak()#text(fill: rgb("',
    muted_c,
    '"), size: 6pt)[',
    pfx,
    ' · ',
    sfx,
    ']],\n'
  ))

  # Col 3: Pop · HH (two lines)
  cat(paste0(
    '  [#text(weight: "semibold", size: 8pt)[',
    fv("pop", 0, m = ","),
    ']',
    '#text(fill: rgb("',
    muted_c,
    '"), size: 6.5pt)[ pop]',
    '#linebreak()#text(weight: "semibold", size: 8pt)[',
    fv("hh", 0, m = ","),
    ']',
    '#text(fill: rgb("',
    muted_c,
    '"), size: 6.5pt)[ hh]],\n'
  ))

  # Col 4: Acres · Dev (two lines)
  cat(paste0(
    '  [#text(weight: "semibold", size: 8pt)[',
    fv("total_acres", 0),
    ']',
    '#text(fill: rgb("',
    muted_c,
    '"), size: 6.5pt)[ ac]',
    '#linebreak()#text(weight: "semibold", size: 8pt)[',
    fv("dev_acres", 0),
    ']',
    '#text(fill: rgb("',
    muted_c,
    '"), size: 6.5pt)[ dev]],\n'
  ))

  # Col 5: OZ 1.0 %
  cat(paste0(
    '  [',
    cell_bar(
      fv("oz1_pct", 1, "%"),
      rel_pct(g("oz1_pct"), col_max$oz1_pct),
      teal_soft
    ),
    '],\n'
  ))

  # Col 6: MUorC % (teal bar)
  cat(paste0(
    '  [',
    cell_bar(
      fv("wc_muc_pct", 1, "%"),
      rel_pct(g("wc_muc_pct"), col_max$muc),
      teal
    ),
    '],\n'
  ))

  # Col 7: ATO /100 (teal-mid bar)
  cat(paste0(
    '  [',
    cell_bar(
      fv("ato_composite", 1),
      rel_pct(g("ato_composite"), col_max$ato),
      teal_mid
    ),
    '],\n'
  ))

  # Col 8: SAP HH/ac (teal-soft bar)
  cat(paste0(
    '  [',
    cell_bar(
      fv("sap_planned_hh_per_acre", 1),
      rel_pct(g("sap_planned_hh_per_acre"), col_max$sap),
      teal_soft
    ),
    '],\n'
  ))

  # Col 9: Growth ERU/ac (amber bar)
  cat(paste0(
    '  [',
    cell_bar(
      fv("eru_per_acre", 1),
      rel_pct(g("eru_per_acre"), col_max$eru),
      amber
    ),
    '],\n'
  ))

  # Col 10: Likelihood pill
  oz_cls_raw <- if (is.na(g("oztoolclassification"))) {
    NA_character_
  } else {
    as.character(g("oztoolclassification"))
  }
  if (is.na(oz_cls_raw)) {
    cat(paste0('  [#text(fill: rgb("', muted_c, '"))[—]],\n'))
  } else {
    cat(paste0(
      '  [#box(fill: rgb("',
      oz_color(oz_cls_raw),
      '"), radius: 3pt, inset: (x: 5pt, y: 3pt))',
      '[#text(fill: white, size: 7pt, weight: "bold")[',
      oz_class_short(oz_cls_raw),
      ']]],\n'
    ))
  }
}

cat(')\n')
cat("```\n")
```
