---
params:
  county: "Salt Lake"
  data_dir: "data/remotes"
format:
  typst:
    margin:
      x: 0.4in
      y: 0.35in
    fontsize: 8pt
---

```{=typst}
#set page(flipped: true)
#set text(font: "Poppins", size: 8pt)
```

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)
library(purrr)
library(yaml)

settings    <- yaml::read_yaml("config/settings.yml")
report_data <- readRDS(file.path(params$data_dir, "report_data.rds"))
tract_maps  <- readRDS(file.path(params$data_dir, "tract_maps.rds"))

county_data <- dplyr::filter(report_data, county == params$county) |>
  dplyr::arrange(GEOID)

county_maps <- tract_maps[county_data$GEOID]
```

```{r}
#| label: save-maps
#| include: false
map_dir <- file.path(params$data_dir, "map_cache")
dir.create(map_dir, showWarnings = FALSE)

map_paths <- purrr::imap_chr(county_maps, function(map, geoid) {
  path <- file.path(map_dir, paste0(geoid, ".png"))
  ggplot2::ggsave(
    filename = path,
    plot     = map,
    width    = settings$map_width_in,
    height   = settings$map_height_in,
    dpi      = 200,
    bg       = "white"
  )
  path
})
```

```{r}
#| label: pages
#| results: asis
#| echo: false

fmt_val <- function(x, digits = 1, suffix = "", big_mark = "") {
  if (length(x) == 0 || is.na(x)) return("—")
  if (is.numeric(x) && !is.finite(x)) return("—")
  rounded <- round(x, digits)
  if (digits == 0 && nchar(big_mark) > 0) {
    return(paste0(format(as.integer(rounded), big.mark = big_mark, scientific = FALSE), suffix))
  }
  paste0(format(rounded, scientific = FALSE), suffix)
}

oz_color <- function(cls) {
  dplyr::case_when(
    grepl("Likely to attract",   cls, ignore.case = TRUE) &
      !grepl("More|Less|Un",     cls, ignore.case = TRUE) ~ "#2a7f46",
    grepl("More likely",         cls, ignore.case = TRUE) ~ "#0072b5",
    grepl("Less likely",         cls, ignore.case = TRUE) ~ "#e07b39",
    grepl("Unlikely|not likely", cls, ignore.case = TRUE) ~ "#c0392b",
    TRUE ~ "#888888"
  )
}

oz_class_short <- function(cls) {
  dplyr::case_when(
    grepl("Likely to attract",   cls, ignore.case = TRUE) &
      !grepl("More|Less|Un",     cls, ignore.case = TRUE) ~ "Likely",
    grepl("More likely",         cls, ignore.case = TRUE) ~ "More Likely",
    grepl("Less likely",         cls, ignore.case = TRUE) ~ "Less Likely",
    grepl("Unlikely|not likely", cls, ignore.case = TRUE) ~ "Unlikely",
    TRUE ~ "Unknown"
  )
}

# ── Page header ──────────────────────────────────────────────────────────────
cat("``` {=typst}\n")
cat(sprintf(
  '#block(width: 100%%, fill: rgb("#003b4f"), inset: (x: 0.15in, y: 0.06in))[#grid(columns: (1fr, auto))[#text(fill: white, weight: "bold", size: 9pt)[%s County — OZ 2.0 Eligible Tracts]][#align(right)[#text(fill: rgb("#7ec8e3"), size: 7pt)[Opportunity Zones Info Sheet | WFRC]]]]\n',
  params$county
))
cat('#line(length: 100%, stroke: (paint: rgb("#7ec8e3"), thickness: 1.5pt))\n')
cat('#v(0.06in)\n')

# ── Summary table ─────────────────────────────────────────────────────────────
# Columns: map | tract | profile | oz1 | wc | ato | sap | eru | oz_likelihood
# Fixed widths sum to 10.20in (usable landscape width with x: 0.4in margins)
cat('#table(\n')
cat('  columns: (1.8in, 0.90in, 1.40in, 0.80in, 1.0in, 0.90in, 1.1in, 1.1in, 1.20in),\n')
cat('  stroke: none,\n')
cat('  fill: (col, row) => if row == 0 { rgb("#003b4f") } else if calc.odd(row) { rgb("#ddeef5") } else { white },\n')
cat('  inset: (x: 5pt, y: 5pt),\n')
cat('  align: (col, row) => if row == 0 { center + horizon } else if col <= 2 { left + top } else if col == 8 { center + top } else { right + top },\n')

cat('  table.header(\n')
header_labels <- c(
  "Map",
  "Tract GEOID",
  "Tract Profile",
  "OZ 1.0#linebreak()Acres / %",
  "Wasatch Choice#linebreak()Center Coverage",
  "Access to#linebreak()Opportunities#linebreak()Score (0–100)",
  "Station Area Plan#linebreak()Housing (HH/Acre)",
  "Projected Growth#linebreak()2027–37 (ERU/Acre)",
  "OZ Investment#linebreak()Likelihood#linebreak()#[(Urban Institute)]"
)
for (lbl in header_labels) {
  cat(sprintf('    [#text(fill: white, weight: "bold", size: 7pt)[%s]],\n', lbl))
}
cat('  ),\n')

# ── Data rows ─────────────────────────────────────────────────────────────────
for (i in seq_len(nrow(county_data))) {
  tract <- county_data[i, ]
  geoid <- as.character(tract$GEOID)
  g     <- function(col) if (col %in% names(tract)) tract[[col]] else NA_real_
  fv    <- function(col, d = 1, s = "", m = "") fmt_val(g(col), d, s, m)

  map_path_fwd <- gsub("\\\\", "/", map_paths[[geoid]])

  # Col 1: Map
  cat(sprintf('  [#image("%s", width: 100%%)],\n', map_path_fwd))

  # Col 2: Tract ID
  cat(sprintf('  [%s],\n', geoid))

  # Col 3: Tract Profile (Pop, HH, Total Acres, Dev. Acres)
  pop_str   <- fv("pop",        0, m = ",")
  hh_str    <- fv("hh",         0, m = ",")
  acres_str <- fv("total_acres", 1)
  devac_str <- fv("dev_acres",  1)
  cat(paste0(
    '  [#strong[Pop.:] ', pop_str,
    ' #h(4pt) #strong[HH:] ', hh_str,
    '#linebreak()#strong[Total Acres:] ', acres_str,
    '#linebreak()#strong[Dev. Acres:] ', devac_str,
    '],\n'
  ))

  # Col 4: OZ 1.0 (acres on line 1, pct on line 2)
  cat(paste0(
    '  [', fv("oz1_acres", 1), ' ac#linebreak()',
    fv("oz1_pct", 1, "%"), '],\n'
  ))

  # Col 5: Wasatch Choice Center Coverage (plain number)
  cat(paste0('  [', fv("wc_muc_pct", 1, "%"), '],\n'))

  # Col 6: Access to Opportunities Score (plain number)
  cat(paste0('  [', fv("ato_composite", 1), '],\n'))

  # Col 7: Station Area Plan Housing (HH/Acre)
  cat(paste0('  [', fv("sap_planned_hh_per_acre", 2), '],\n'))

  # Col 8: Projected Growth 2027-37 (ERU/Acre)
  cat(paste0('  [', fv("eru_per_acre", 2), '],\n'))

  # Col 9: OZ Investment Likelihood — last column, attributed to Urban Institute
  oz_cls_raw <- if (is.na(g("oztoolclassification"))) NA_character_
                else as.character(g("oztoolclassification"))
  if (is.na(oz_cls_raw)) {
    cat('  [—],\n')
  } else {
    cat(paste0(
      '  [#box(fill: rgb("', oz_color(oz_cls_raw), '"), radius: 2pt, inset: (x: 4pt, y: 3pt))',
      '[#text(fill: white, size: 7pt, weight: "bold")[', oz_class_short(oz_cls_raw), ']]],\n'
    ))
  }
}

cat(')\n')
cat('#v(0.04in)\n')
cat('#line(length: 100%, stroke: (paint: rgb("#dddddd"), thickness: 0.5pt))\n')
cat('#v(1pt)\n')
cat('#align(right)[#text(size: 5.5pt, fill: rgb("#888888"))[Sources: Federal OZ 2.0 eligibility, Urban Institute OZ Tool, WFRC/MAG regional data. Growth projections 2027–2037. ATO = Access to Opportunities. ERU = Equivalent Residential Unit. SAP = Station Area Plan. WC = Wasatch Choice.]]\n')
cat("```\n")
```
