---
title: "OZ 2.0 Candidate Tracts — Land Use & Transportation Info Sheets"
subtitle: "WFRC & MAG Region"
format: html
---

# Load Libraries

```{r}
#| label: libraries
#| include: false
# Data access
library(yaml)
library(arcgislayers)
library(tigris)
library(tidycensus)

# Spatial
library(sf)

# Data wrangling
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

# Mapping and output
library(ggplot2)
library(knitr)

# Report assembly
library(quarto)
library(tinytable)

# Options
options(tigris_use_cache = FALSE) # caching handled by src/download_data.R
options(scipen = 999) # suppress scientific notation
```

# Data

## Prepare

```{r}
#| label: prepare
source("src/download_data.R")
config <- load_config()
settings <- load_settings()

crs <- settings$crs_analysis
remotes_dir <- config$data_dir_remotes
map_width_in <- settings$map_width_in
map_height_in <- settings$map_height_in
knitr::opts_chunk$set(fig.width = map_width_in, fig.height = map_height_in)

# County names in report order — WFRC region first, then MAG
county_order <- c(
  "Box Elder",
  "Weber",
  "Davis",
  "Salt Lake",
  "Tooele",
  "Morgan", # WFRC
  "Utah",
  "Summit",
  "Wasatch" # MAG
)

# Resolve FIPS codes dynamically via tidycensus internals
state_fips <- tidycensus:::validate_state(state = settings$state)
county_fips <- vapply(
  county_order,
  function(x) {
    tidycensus:::validate_county(state = settings$state, county = x)
  },
  character(1)
)
# county_fips: named vector — names = county names, values = 3-digit FIPS
# county_geoids: 5-digit state+county FIPS prefix used for ACS GEOID filtering
county_geoids <- paste0(state_fips, county_fips)

# Ensure all layers are cached; returns named list of file paths
cache_paths <- download_all(config)

# Shared SQL IN clauses for county filtering
county_fips_in <- paste0("'", county_fips, "'", collapse = ", ")
county_names_in <- paste0("'", county_order, "'", collapse = ", ")

# TAZ projection layers use integer CO_FIPS (standard 3-digit FIPS as integer: '035' -> 35)
co_fips_int_in <- paste(as.integer(county_fips), collapse = ", ")
```

## Load

```{r}
#| label: load
oz_tracts <- sf::read_sf(
  cache_paths$oz_eligible_tracts,
  query = sprintf(
    "SELECT * FROM oz_eligible_tracts
     WHERE oz_lgbl = 'OZ eligible'
       AND STATEFP = '%s'
       AND COUNTYF IN (%s)",
    state_fips,
    county_fips_in
  )
) |>
  sf::st_transform(crs) |>
  dplyr::mutate(
    county = factor(
      names(county_fips)[match(COUNTYF, county_fips)],
      levels = county_order
    ),
    mfi = dplyr::na_if(mfi, "not available"),
    mfi = as.integer(readr::parse_number(mfi))
  )

# Urban Institute layer — attributes only; filtered to candidate tract GEOIDs
oz_ui <- sf::read_sf(
  cache_paths$oz_ui_analysis,
  query = sprintf(
    "SELECT * FROM oz_ui_analysis WHERE GEOID20 IN (%s)",
    paste0("'", oz_tracts$GEOID, "'", collapse = ", ")
  )
) |>
  sf::st_drop_geometry()

counties <- sf::read_sf(
  cache_paths$counties,
  query = sprintf(
    "SELECT COUNTYFP, NAME, geom FROM counties WHERE COUNTYFP IN (%s)",
    county_fips_in
  )
) |>
  sf::st_transform(crs)

tracts <- sf::read_sf(
  cache_paths$tracts,
  query = sprintf(
    "SELECT GEOID, COUNTYFP, geom FROM tracts WHERE COUNTYFP IN (%s)",
    county_fips_in
  )
) |>
  sf::st_transform(crs)

# Places: no COUNTYFP field — combine spatial filter (oz_tracts bbox in WGS84)
# with column selection. wkt_filter applies before query in OGR.
oz_bbox_wgs84 <- oz_tracts |>
  sf::st_transform(4326) |>
  sf::st_bbox() |>
  sf::st_as_sfc() |>
  sf::st_as_text()

places <- sf::read_sf(
  cache_paths$places,
  query = "SELECT NAME, geom FROM places",
  wkt_filter = oz_bbox_wgs84
) |>
  sf::st_transform(crs)

# Filter to study counties at read time; STATENAME and COUNTYNAME are the available attribute filters
oz_1 <- sf::read_sf(
  cache_paths$oz_1_tracts,
  query = sprintf(
    "SELECT OBJECTID, geom FROM oz_1_tracts
     WHERE STATENAME = '%s'
       AND COUNTYNAME IN (%s)",
    settings$state_name,
    county_names_in
  )
) |>
  sf::st_transform(crs)

# CenterName included for optional tract labelling
wc_centers <- sf::read_sf(
  cache_paths$wasatch_choice_centers,
  query = "SELECT OBJECTID, CenterName, CenterType, geom FROM wasatch_choice_centers"
) |>
  sf::st_transform(crs) |>
  sf::st_make_valid()

uta_stops <- sf::read_sf(
  cache_paths$uta_stops,
  query = "SELECT OBJECTID, Station, geom FROM uta_stops"
) |>
  sf::st_transform(crs)

# Station + Line needed for crosswalk join; OBJECTID kept for sap_overlap block below
sap_buffers <- sf::read_sf(
  cache_paths$sap_buffers,
  query = "SELECT OBJECTID, Station, Line, geom FROM sap_buffers"
) |>
  sf::st_transform(crs) |>
  sf::st_make_valid()

# Geometry only — xid is the OID field for this layer
freeway_exits <- sf::read_sf(
  cache_paths$freeway_interchanges,
  query = "SELECT xid, geom FROM freeway_interchanges"
) |>
  sf::st_transform(crs)

# TYPE / SUBTYPE classify unit type (SFD, MF, etc.); UNIT_COUNT and ACRES drive metrics
housing_inv <- sf::read_sf(
  cache_paths$housing_inventory,
  query = "SELECT OBJECTID, TYPE, SUBTYPE, UNIT_COUNT, ACRES, geom FROM housing_inventory"
) |>
  sf::st_transform(crs)

ato <- sf::read_sf(
  cache_paths$ato,
  query = sprintf(
    "SELECT TAZID, CO_FIPS, DEVACRES,
            JOBAUTO_23, HHAUTO_23, JOBTRANSIT_23, HHTRANSIT_23, geom
     FROM ato WHERE CO_FIPS IN (%s)",
    co_fips_int_in
  )
) |>
  sf::st_transform(crs)

taz_hh_projections <- sf::read_sf(
  cache_paths$taz_hh_projections,
  query = sprintf(
    "SELECT TAZID, YEAR2027, YEAR2037, geom FROM taz_hh_projections WHERE CO_FIPS IN (%s)",
    co_fips_int_in
  )
) |>
  sf::st_transform(crs)

taz_job_projections <- sf::read_sf(
  cache_paths$taz_job_projections,
  query = sprintf(
    "SELECT TAZID, YEAR2027, YEAR2037, geom FROM taz_job_projections WHERE CO_FIPS IN (%s)",
    co_fips_int_in
  )
) |>
  sf::st_transform(crs)

taz_pop_projections <- sf::read_sf(
  cache_paths$taz_pop_projections,
  query = sprintf(
    "SELECT TAZID, YEAR2027, YEAR2037, geom FROM taz_pop_projections WHERE CO_FIPS IN (%s)",
    co_fips_int_in
  )
) |>
  sf::st_transform(crs)

# --- ACS: non-spatial, read from RDS; pivot to one row per tract ---

acs <- readr::read_csv(
  cache_paths$acs,
  col_types = readr::cols(GEOID = readr::col_character()),
  show_col_types = FALSE
) |>
  dplyr::filter(stringr::str_sub(GEOID, 1, 5) %in% county_geoids) |>
  dplyr::select(GEOID, variable, estimate) |>
  tidyr::pivot_wider(names_from = variable, values_from = estimate)
```

# Tract Info

```{r}
#| label: tract-info
# Total tract area (sqm → acres; CRS 26912 is metric)
total_acres_df <- data.frame(
  GEOID = oz_tracts$GEOID,
  total_acres = as.numeric(sf::st_area(oz_tracts)) / 4046.8564
)

# OZ 1.0 overlap — oz_1 already filtered to state at read time
oz1_overlap <- suppressWarnings(
  sf::st_intersection(
    oz_tracts |> dplyr::select(GEOID),
    oz_1 |> dplyr::select(OBJECTID)
  )
) |>
  dplyr::mutate(overlap_m2 = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(overlap_m2 > 0) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(oz1_acres = sum(overlap_m2) / 4046.8564, .groups = "drop")

# Developable acres — area-proportioned from ATO DEVACRES where TAZ boundaries
# cross tract boundaries: contribution = DEVACRES × (intersection_area / TAZ_area)
ato_areas_df <- data.frame(
  TAZID = ato$TAZID,
  taz_area = as.numeric(sf::st_area(ato))
)
dev_acres_df <- suppressWarnings(
  sf::st_intersection(
    oz_tracts |> dplyr::select(GEOID),
    ato |> dplyr::select(TAZID, DEVACRES)
  )
) |>
  dplyr::mutate(int_area = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(int_area > 0) |>
  dplyr::left_join(ato_areas_df, by = "TAZID") |>
  dplyr::mutate(dev_acres_contrib = DEVACRES * (int_area / taz_area)) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(dev_acres = sum(dev_acres_contrib), .groups = "drop")

# Assemble: one row per OZ 2.0 candidate tract, sorted by county then tract ID
tract_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID, county, COUNTYF) |>
  dplyr::left_join(
    dplyr::select(acs, GEOID, pop = population, hh = households),
    by = "GEOID"
  ) |>
  dplyr::left_join(total_acres_df, by = "GEOID") |>
  dplyr::left_join(oz1_overlap, by = "GEOID") |>
  dplyr::mutate(
    oz1_acres = dplyr::coalesce(oz1_acres, 0),
    oz1_pct = round(oz1_acres / total_acres * 100, 1)
  ) |>
  dplyr::left_join(dev_acres_df, by = "GEOID") |>
  dplyr::arrange(county, GEOID)
```

## Centers

```{r}
#| label: centers
# Intersect oz_tracts with Wasatch Choice centers — acres per CenterType per tract
wc_intersection <- suppressWarnings(
  sf::st_intersection(
    dplyr::select(oz_tracts, GEOID),
    dplyr::select(wc_centers, CenterType)
  )
) |>
  dplyr::mutate(int_m2 = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(int_m2 > 0) |>
  dplyr::group_by(GEOID, CenterType) |>
  dplyr::summarize(acres = sum(int_m2) / 4046.8564, .groups = "drop")

# Pivot wide: one wc_acres_* and one wc_pct_* column per CenterType.
# col_id converts "Metropolitan Center" → "metropolitan_center" etc.
centers_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID) |>
  dplyr::left_join(wc_intersection, by = "GEOID") |>
  dplyr::left_join(total_acres_df, by = "GEOID") |>
  dplyr::mutate(
    pct = round(acres / total_acres * 100, 1),
    col_id = tolower(gsub("[^a-zA-Z0-9]+", "_", CenterType))
  ) |>
  dplyr::select(GEOID, col_id, acres, pct) |>
  tidyr::pivot_wider(
    names_from = col_id,
    values_from = c(acres, pct),
    values_fill = 0,
    names_glue = "wc_{.value}_{col_id}"
  )
```

```{r}
#| label: centers-composite
# 1) % of tract covered by MUorC (Metropolitan, Urban, or City) center types.
#    Center polygons do not overlap within a tract, so summing individual pcts is valid.
muc_pct_cols <- paste0(
  "wc_pct_",
  c("metropolitan_center", "urban_center", "city_center")
)
centers_info <- centers_info |>
  dplyr::mutate(
    wc_muc_pct = rowSums(
      dplyr::pick(dplyr::any_of(muc_pct_cols)),
      na.rm = TRUE
    )
  )
```

## Housing

```{r}
#| label: housing
# Assign each housing polygon to the tract containing its centroid.
# Centroid assignment is appropriate for parcel-level inventory data — units
# belong to one tract; area-proportioning is not needed.
housing_joined <- sf::st_join(
  sf::st_centroid(housing_inv),
  dplyr::select(oz_tracts, GEOID),
  join = sf::st_within
) |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(GEOID))

# Totals per tract
housing_totals <- housing_joined |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(
    units_total = sum(UNIT_COUNT, na.rm = TRUE),
    res_acres_total = sum(ACRES, na.rm = TRUE),
    .groups = "drop"
  )

# Unit counts by TYPE — pivot wide; column names reflect actual TYPE values
# in the data, so verify TYPE values before interpreting columns
housing_by_type <- housing_joined |>
  dplyr::group_by(GEOID, TYPE) |>
  dplyr::summarize(units = sum(UNIT_COUNT, na.rm = TRUE), .groups = "drop") |>
  tidyr::pivot_wider(
    names_from = TYPE,
    values_from = units,
    values_fill = 0L,
    names_prefix = "units_"
  )

housing_acres <- housing_joined |>
  dplyr::mutate(
    acre_cat = dplyr::case_when(
      TYPE == "single_family" ~ "sfd",
      TYPE == "multi_family" ~ "mf_sfa",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::filter(!is.na(acre_cat)) |>
  dplyr::group_by(GEOID, acre_cat) |>
  dplyr::summarize(acres = sum(ACRES, na.rm = TRUE), .groups = "drop") |>
  tidyr::complete(
    GEOID,
    acre_cat = c("sfd", "mf_sfa"),
    fill = list(acres = 0)
  ) |>
  tidyr::pivot_wider(
    names_from = acre_cat,
    values_from = acres,
    values_fill = 0,
    names_prefix = "res_acres_"
  )

housing_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID) |>
  dplyr::left_join(housing_totals, by = "GEOID") |>
  dplyr::left_join(housing_by_type, by = "GEOID") |>
  dplyr::left_join(housing_acres, by = "GEOID") |>
  dplyr::mutate(dplyr::across(where(is.numeric), ~ dplyr::coalesce(.x, 0)))
```

## Transit

```{r}
#| label: transit
# ---- SAP station points within each tract ----
stops_within <- sf::st_join(
  dplyr::select(oz_tracts, GEOID),
  dplyr::select(uta_stops, Station),
  join = sf::st_contains
) |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(Station)) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(
    transit_stop_count = dplyr::n(),
    transit_stop_names = paste(sort(unique(stopname)), collapse = "; "),
    .groups = "drop"
  )

# Distance to nearest stop — computed for all tracts via spatial index;
# set to NA for tracts that already have stops within
nearest_stop_idx <- sf::st_nearest_feature(oz_tracts, uta_stops)
transit_dist_df <- data.frame(
  GEOID = oz_tracts$GEOID,
  dist_edge_mi = as.numeric(sf::st_distance(
    oz_tracts,
    uta_stops[nearest_stop_idx, ],
    by_element = TRUE
  )) *
    0.000621371,
  dist_centroid_mi = as.numeric(sf::st_distance(
    suppressWarnings(sf::st_centroid(oz_tracts)),
    uta_stops[nearest_stop_idx, ],
    by_element = TRUE
  )) *
    0.000621371
)

# ---- SAP buffer intersection — acres and % of tract ----
sap_overlap <- suppressWarnings(
  sf::st_intersection(
    dplyr::select(oz_tracts, GEOID),
    dplyr::select(sap_buffers, OBJECTID)
  )
) |>
  dplyr::mutate(sap_m2 = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(sap_m2 > 0) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(sap_acres = sum(sap_m2) / 4046.8564, .groups = "drop")

# ---- ATO transit scores — weighted avg by proportioned DEVACRES ----
# Reuses ato_areas_df (TAZ total areas) computed in tract-info chunk
ato_transit_df <- suppressWarnings(
  sf::st_intersection(
    dplyr::select(oz_tracts, GEOID),
    dplyr::select(ato, TAZID, DEVACRES, JOBTRANSIT_23, HHTRANSIT_23)
  )
) |>
  dplyr::mutate(int_area = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(int_area > 0) |>
  dplyr::left_join(ato_areas_df, by = "TAZID") |>
  dplyr::mutate(weight = DEVACRES * (int_area / taz_area)) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(
    ato_jobtransit = sum(JOBTRANSIT_23 * weight, na.rm = TRUE) /
      sum(weight, na.rm = TRUE),
    ato_hhtransit = sum(HHTRANSIT_23 * weight, na.rm = TRUE) /
      sum(weight, na.rm = TRUE),
    .groups = "drop"
  )

# ---- Assemble ----
# Note: planned Phase 1 station flags are manual annotations (no layer available);
# leave as a join target for later when that data is provided.
transit_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID) |>
  dplyr::left_join(stops_within, by = "GEOID") |>
  dplyr::mutate(transit_stop_count = dplyr::coalesce(transit_stop_count, 0L)) |>
  dplyr::left_join(transit_dist_df, by = "GEOID") |>
  dplyr::mutate(
    dist_edge_mi = dplyr::if_else(
      transit_stop_count > 0,
      NA_real_,
      dist_edge_mi
    ),
    dist_centroid_mi = dplyr::if_else(
      transit_stop_count > 0,
      NA_real_,
      dist_centroid_mi
    )
  ) |>
  dplyr::left_join(sap_overlap, by = "GEOID") |>
  dplyr::left_join(
    dplyr::select(total_acres_df, GEOID, total_acres),
    by = "GEOID"
  ) |>
  dplyr::mutate(
    sap_acres = dplyr::coalesce(sap_acres, 0),
    sap_pct = round(sap_acres / total_acres * 100, 1)
  ) |>
  dplyr::select(-total_acres) |>
  dplyr::left_join(ato_transit_df, by = "GEOID")
```

```{r}
#| label: transit-sap-housing
# Planned housing units in SAP areas per tract, normalized by SAP overlap acres.
# Source: data/SAP Housing and Connections Metrics.xlsx (Byron Head, 2025).
# Crosswalk maps GPKG (Station, Line) → Excel group + housing_units.
# City-boundary polygon splits within the same group are adjacent (non-overlapping),
# so summing individual polygon areas gives the correct group total without dissolving.
sap_crosswalk <- readr::read_csv(
  "data/sap_housing_crosswalk.csv",
  comment = "#",
  col_types = readr::cols_only(
    station_gpkg = readr::col_character(),
    line_gpkg = readr::col_character(),
    group_id = readr::col_character(),
    housing_units = readr::col_double()
  )
)

# Per-polygon area → group total area → hh_density for each group
sap_group_density <- sap_buffers |>
  dplyr::inner_join(
    sap_crosswalk,
    by = c("Station" = "station_gpkg", "Line" = "line_gpkg")
  ) |>
  dplyr::mutate(poly_area_acres = as.numeric(sf::st_area(geom)) / 4046.8564) |>
  sf::st_drop_geometry() |>
  dplyr::group_by(group_id, housing_units) |>
  dplyr::summarize(group_area_acres = sum(poly_area_acres), .groups = "drop") |>
  dplyr::mutate(
    hh_density = dplyr::if_else(
      group_area_acres > 0,
      housing_units / group_area_acres,
      0
    )
  )

# Attach group density back to individual (city-split) polygons for intersection
sap_with_density <- sap_buffers |>
  dplyr::inner_join(
    sap_crosswalk,
    by = c("Station" = "station_gpkg", "Line" = "line_gpkg")
  ) |>
  dplyr::left_join(
    dplyr::select(sap_group_density, group_id, hh_density),
    by = "group_id"
  )

# Intersect raw polygons with OZ tracts; area-weighted average density per tract
sap_housing_overlap <- suppressWarnings(
  sf::st_intersection(
    dplyr::select(oz_tracts, GEOID),
    dplyr::select(sap_with_density, group_id, hh_density)
  )
) |>
  dplyr::mutate(overlap_acres = as.numeric(sf::st_area(geom)) / 4046.8564) |>
  sf::st_drop_geometry() |>
  dplyr::filter(overlap_acres > 0) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(
    sap_planned_hh_per_acre = sum(overlap_acres * hh_density) /
      sum(overlap_acres),
    .groups = "drop"
  )

transit_info <- transit_info |>
  dplyr::left_join(sap_housing_overlap, by = "GEOID") |>
  dplyr::mutate(
    sap_planned_hh_per_acre = dplyr::coalesce(sap_planned_hh_per_acre, 0)
  )
```

## Auto

```{r}
#| label: auto
# ---- Freeway interchanges within each tract ----
interchanges_within <- sf::st_join(
  dplyr::select(oz_tracts, GEOID),
  dplyr::select(freeway_exits, xid),
  join = sf::st_contains
) |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(xid)) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(freeway_exit_count = dplyr::n(), .groups = "drop")

# Distance to nearest interchange — cleared for tracts that already have one
nearest_exit_idx <- sf::st_nearest_feature(oz_tracts, freeway_exits)
exit_dist_df <- data.frame(
  GEOID = oz_tracts$GEOID,
  exit_dist_edge_mi = as.numeric(sf::st_distance(
    oz_tracts,
    freeway_exits[nearest_exit_idx, ],
    by_element = TRUE
  )) *
    0.000621371,
  exit_dist_centroid_mi = as.numeric(sf::st_distance(
    suppressWarnings(sf::st_centroid(oz_tracts)),
    freeway_exits[nearest_exit_idx, ],
    by_element = TRUE
  )) *
    0.000621371
)

# ---- ATO auto scores — weighted avg by proportioned DEVACRES ----
# Reuses ato_areas_df (TAZ total areas) computed in tract-info chunk
ato_auto_df <- suppressWarnings(
  sf::st_intersection(
    dplyr::select(oz_tracts, GEOID),
    dplyr::select(ato, TAZID, DEVACRES, JOBAUTO_23, HHAUTO_23)
  )
) |>
  dplyr::mutate(int_area = as.numeric(sf::st_area(geom))) |>
  sf::st_drop_geometry() |>
  dplyr::filter(int_area > 0) |>
  dplyr::left_join(ato_areas_df, by = "TAZID") |>
  dplyr::mutate(weight = DEVACRES * (int_area / taz_area)) |>
  dplyr::group_by(GEOID) |>
  dplyr::summarize(
    ato_jobauto = sum(JOBAUTO_23 * weight, na.rm = TRUE) /
      sum(weight, na.rm = TRUE),
    ato_hhauto = sum(HHAUTO_23 * weight, na.rm = TRUE) /
      sum(weight, na.rm = TRUE),
    .groups = "drop"
  )

# ---- Assemble ----
auto_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID) |>
  dplyr::left_join(interchanges_within, by = "GEOID") |>
  dplyr::mutate(freeway_exit_count = dplyr::coalesce(freeway_exit_count, 0L)) |>
  dplyr::left_join(exit_dist_df, by = "GEOID") |>
  dplyr::mutate(
    exit_dist_edge_mi = dplyr::if_else(
      freeway_exit_count > 0,
      NA_real_,
      exit_dist_edge_mi
    ),
    exit_dist_centroid_mi = dplyr::if_else(
      freeway_exit_count > 0,
      NA_real_,
      exit_dist_centroid_mi
    )
  ) |>
  dplyr::left_join(ato_auto_df, by = "GEOID")
```

```{r}
#| label: ato-composite
# 2) Composite ATO score — "both modes both ways":
#    min-max normalize each of the 4 components to 0–100 across all tracts,
#    then average. Yields a single 0–100 index for cross-tract comparison.
ato_composite_df <- transit_info |>
  dplyr::select(GEOID, ato_jobtransit, ato_hhtransit) |>
  dplyr::left_join(
    dplyr::select(auto_info, GEOID, ato_jobauto, ato_hhauto),
    by = "GEOID"
  ) |>
  dplyr::mutate(
    dplyr::across(
      c(ato_jobtransit, ato_hhtransit, ato_jobauto, ato_hhauto),
      ~ {
        mn <- min(.x, na.rm = TRUE)
        mx <- max(.x, na.rm = TRUE)
        if (mx > mn) (.x - mn) / (mx - mn) * 100 else 0
      },
      .names = "{.col}_norm"
    ),
    ato_composite = rowMeans(
      dplyr::pick(dplyr::ends_with("_norm")),
      na.rm = TRUE
    )
  ) |>
  dplyr::select(GEOID, ato_composite)
```

## Projected Growth

```{r}
#| label: growth
# Helper: area-proportion (YEAR2037 - YEAR2027) from a TAZ projection layer
# onto oz_tracts. Uses plain area proportioning — no DEVACRES in these layers.
.proportion_growth <- function(taz_layer, value_col) {
  taz_areas <- data.frame(
    TAZID = taz_layer$TAZID,
    taz_area = as.numeric(sf::st_area(taz_layer))
  )
  suppressWarnings(
    sf::st_intersection(
      dplyr::select(oz_tracts, GEOID),
      dplyr::select(taz_layer, TAZID, YEAR2027, YEAR2037)
    )
  ) |>
    dplyr::mutate(int_area = as.numeric(sf::st_area(geom))) |>
    sf::st_drop_geometry() |>
    dplyr::filter(int_area > 0) |>
    dplyr::left_join(taz_areas, by = "TAZID") |>
    dplyr::mutate(contrib = (YEAR2037 - YEAR2027) * (int_area / taz_area)) |>
    dplyr::group_by(GEOID) |>
    dplyr::summarize(
      !!value_col := round(sum(contrib, na.rm = TRUE)),
      .groups = "drop"
    )
}

hh_growth_df <- .proportion_growth(taz_hh_projections, "hh_added")
pop_growth_df <- .proportion_growth(taz_pop_projections, "pop_added")
job_growth_df <- .proportion_growth(taz_job_projections, "jobs_added")

growth_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID) |>
  dplyr::left_join(hh_growth_df, by = "GEOID") |>
  dplyr::left_join(pop_growth_df, by = "GEOID") |>
  dplyr::left_join(job_growth_df, by = "GEOID") |>
  dplyr::mutate(dplyr::across(where(is.numeric), ~ dplyr::coalesce(.x, 0L)))
```

```{r}
#| label: growth-eru
# 4) Equivalent Residential Units (ERU) added, normalized by total tract acres.
#    Converts projected HH and job growth to a common density metric so tracts
#    can be compared regardless of size. Conversion: 1 HH = 1 ERU, 1 job = 0.4 ERU
#    (common planning assumption; adjust eru_jobs_factor if a different standard applies).
eru_jobs_factor <- 0.4

growth_info <- growth_info |>
  dplyr::left_join(
    dplyr::select(total_acres_df, GEOID, total_acres),
    by = "GEOID"
  ) |>
  dplyr::mutate(
    eru_added = hh_added + jobs_added * eru_jobs_factor,
    eru_per_acre = dplyr::if_else(
      total_acres > 0,
      round(eru_added / total_acres, 2),
      NA_real_
    )
  ) |>
  dplyr::select(-total_acres)
```

## Urban Institute

```{r}
#| label: urban-institute
# Socioeconomic indicators from oz_eligible_tracts (pvrty_r, unmpl_r, mfi) joined
# with oztoolclassification from the Urban Institute layer.
# Inspect names(oz_ui) to see all additional fields available from that layer.
ui_info <- oz_tracts |>
  sf::st_drop_geometry() |>
  dplyr::select(GEOID, pvrty_r, unmpl_r, mfi) |>
  dplyr::left_join(
    oz_ui |>
      dplyr::select(GEOID20, oztoolclassification) |>
      dplyr::rename(GEOID = GEOID20),
    by = "GEOID"
  )
```

# Prepare Maps

```{r}
#| label: maps
# Expand bbox by a scale factor from its center, preserving CRS class/attributes
.expand_bbox <- function(bbox, factor) {
  cx <- (bbox["xmin"] + bbox["xmax"]) / 2
  cy <- (bbox["ymin"] + bbox["ymax"]) / 2
  hw <- (bbox["xmax"] - bbox["xmin"]) / 2 * factor
  hh <- (bbox["ymax"] - bbox["ymin"]) / 2 * factor
  bbox["xmin"] <- cx - hw
  bbox["xmax"] <- cx + hw
  bbox["ymin"] <- cy - hh
  bbox["ymax"] <- cy + hh
  bbox
}

# Pad bbox so its width:height matches the target figure ratio, preventing dead
# space when coord_sf enforces the geographic aspect ratio of the extent.
.enforce_aspect <- function(bbox, width, height) {
  target <- width / height
  bw <- bbox["xmax"] - bbox["xmin"]
  bh <- bbox["ymax"] - bbox["ymin"]
  cx <- (bbox["xmin"] + bbox["xmax"]) / 2
  cy <- (bbox["ymin"] + bbox["ymax"]) / 2
  if ((bw / bh) > target) {
    half <- bw / target / 2
    bbox["ymin"] <- cy - half
    bbox["ymax"] <- cy + half
  } else {
    half <- bh * target / 2
    bbox["xmin"] <- cx - half
    bbox["xmax"] <- cx + half
  }
  bbox
}

# Wasatch Choice center/district type fill colors — kept as-is per spec
wc_center_colors <- c(
  "Metropolitan Center" = "#bc5f8c",
  "Urban Center" = "#ec8369",
  "City Center" = "#f5b86e",
  "Neighborhood Center" = "#fae55c",
  "Education District" = "#cbe3e1",
  "Employment District" = "#f6c4da",
  "Industrial District" = "#eadbf4",
  "Retail District" = "#f8dddf",
  "Special District" = "#d3d3d3"
)

# Warm parchment — map canvas and plot background
map_bg <- "#F7F4EF"

make_tract_map <- function(geoid) {
  focal <- dplyr::filter(oz_tracts, GEOID == geoid)
  other_oz <- dplyr::filter(oz_tracts, GEOID != geoid)
  exp_bb <- .enforce_aspect(
    .expand_bbox(sf::st_bbox(focal), factor = 1.2),
    map_width_in,
    map_height_in
  )
  view <- sf::st_set_crs(sf::st_as_sfc(exp_bb), sf::st_crs(crs))

  ggplot2::ggplot() +
    # 1. Background census tracts — warm light gray, very recessive
    ggplot2::geom_sf(
      data = sf::st_filter(tracts, view),
      fill = "#EDE8E1",
      color = "#D8D2C8",
      linewidth = 0.1
    ) +
    # 2. OZ 1.0 tracts — warm rose-gray underlay shows prior designation
    ggplot2::geom_sf(
      data = sf::st_filter(oz_1, view),
      fill = "#D4C8C4",
      color = NA,
      alpha = 0.5
    ) +
    # 3. Wasatch Choice centers and districts
    ggplot2::geom_sf(
      data = sf::st_filter(wc_centers, view),
      ggplot2::aes(fill = CenterType),
      color = NA,
      alpha = 0.75
    ) +
    ggplot2::scale_fill_manual(
      values = wc_center_colors,
      name = NULL,
      na.value = "#C8C0C0",
      drop = TRUE
    ) +
    # 4. Other OZ 2.0 candidate tracts — pale straw with amber border
    ggplot2::geom_sf(
      data = sf::st_filter(other_oz, view),
      fill = "#F2E8C4",
      color = "#C4A43A",
      linewidth = 0.25,
      alpha = 0.5
    ) +
    # 5. SAP buffers — cool blue over warm palette; temperature contrast reads clearly
    ggplot2::geom_sf(
      data = sf::st_filter(sap_buffers, view),
      fill = "#B8D4E8",
      color = "#6FA4C4",
      linewidth = 0.25,
      alpha = 0.45
    ) +
    # 6. County boundary — thin charcoal, crisp orientation anchor
    ggplot2::geom_sf(
      data = sf::st_filter(counties, view),
      fill = NA,
      color = "#5A5A5A",
      linewidth = 0.35
    ) +
    # 7. Freeway interchanges — solid dark-slate triangle; no white halo at this size
    ggplot2::geom_sf(
      data = sf::st_filter(freeway_exits, view),
      shape = 17,
      color = "#3A4A5A",
      size = 2.0
    ) +
    # 8. UTA Rail / BRT stops — deep burgundy; white stroke separates from dark fills
    ggplot2::geom_sf(
      data = sf::st_filter(uta_stops, view),
      shape = 21,
      fill = "#943030",
      color = "white",
      size = 2.6,
      stroke = 0.4
    ) +
    # 9. Focal tract — opaque charcoal border; fill alpha via scales::alpha() so
    #    only the fill is translucent, leaving the stroke at full opacity
    ggplot2::geom_sf(
      data = focal,
      fill = scales::alpha("#FFFFF0", 0.25),
      color = "#1A1A1A",
      linewidth = 1.2
    ) +
    ggplot2::coord_sf(
      xlim = c(exp_bb["xmin"], exp_bb["xmax"]),
      ylim = c(exp_bb["ymin"], exp_bb["ymax"]),
      crs = sf::st_crs(crs),
      default_crs = sf::st_crs(crs),
      datum = sf::st_crs(crs),
      expand = FALSE
    ) +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "none",
      plot.background = ggplot2::element_rect(fill = map_bg, color = NA),
      plot.margin = ggplot2::margin(0, 0, 0, 0)
    )
}

tract_maps <- purrr::set_names(
  purrr::map(oz_tracts$GEOID, make_tract_map),
  oz_tracts$GEOID
)
```

```{r}
#| eval: false
tract_maps[["49035102900"]]
```

# Assemble Report

Each chunk below has `#| eval: false` and can be run independently in RStudio.
Chunks 2 and 3 load from the RDS files on disk when `report_data` / `tract_maps`
are not already in the session, so you can re-export without re-running the full
data pipeline.

## 1 — Build Master Table

Run this first (or whenever source data changes). Joins all metric tables and
persists `report_data.rds` + `tract_maps.rds` so the export chunks can load them.

```{r}
#| label: report-build
#| eval: false
report_data <- tract_info |>
  dplyr::left_join(centers_info, by = "GEOID") |>
  dplyr::left_join(housing_info, by = "GEOID") |>
  dplyr::left_join(transit_info, by = "GEOID") |>
  dplyr::left_join(auto_info, by = "GEOID") |>
  dplyr::left_join(ato_composite_df, by = "GEOID") |>
  dplyr::left_join(growth_info, by = "GEOID") |>
  dplyr::left_join(ui_info, by = "GEOID")

saveRDS(report_data, file.path(remotes_dir, "report_data.rds"))
saveRDS(tract_maps, file.path(remotes_dir, "tract_maps.rds"))
message("Saved report_data.rds and tract_maps.rds to ", remotes_dir)
```

## 2 — Export Excel

```{r}
#| label: report-excel
#| eval: false
if (!exists("report_data")) {
  report_data <- readRDS(file.path(remotes_dir, "report_data.rds"))
}
if (!exists("tract_maps")) {
  tract_maps <- readRDS(file.path(remotes_dir, "tract_maps.rds"))
}

source("src/export_excel.R")
export_excel(
  report_data = report_data,
  tract_maps = tract_maps,
  settings = settings,
  remotes_dir = remotes_dir
)
```

## 3 — Export PDFs

Summary and detail PDFs are rendered back-to-back per county — one pass through the
county list produces both documents, which avoids reloading the RDS in child sessions
twice per county.

```{r}
#| label: report-pdfs
#| eval: false
if (!exists("report_data")) {
  report_data <- readRDS(file.path(remotes_dir, "report_data.rds"))
}

county_names <- levels(droplevels(report_data$county))
summary_dir <- file.path(settings$output_dir, "summary")
detail_dir <- file.path(settings$output_dir, "detail")
dir.create(summary_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(detail_dir, showWarnings = FALSE, recursive = TRUE)

# Pre-flight: Typst always writes to <input-basename>.pdf before renaming to
# output_file. If a PDF viewer has those files open on Windows the write is
# blocked — clear them now so the pipeline doesn't fail mid-run.
for (.f in c("_county_summary.pdf", "_county_detail.pdf")) {
  if (file.exists(.f) && !suppressWarnings(file.remove(.f))) {
    stop(
      sprintf("Close '%s' in your PDF viewer before running.", .f),
      call. = FALSE
    )
  }
}
rm(.f)

purrr::walk(county_names, function(county_name) {
  slug <- gsub(" ", "_", county_name)

  summary_file <- paste0(slug, "_OZ_Summary.pdf")
  quarto::quarto_render(
    input = "_county_summary.qmd",
    output_file = summary_file,
    execute_params = list(county = county_name, data_dir = remotes_dir)
  )
  file.rename(summary_file, file.path(summary_dir, summary_file))

  detail_file <- paste0(slug, "_OZ_Detail.pdf")
  quarto::quarto_render(
    input = "_county_detail.qmd",
    output_file = detail_file,
    execute_params = list(county = county_name, data_dir = remotes_dir)
  )
  file.rename(detail_file, file.path(detail_dir, detail_file))
})
```
